### 2018-12-23 Partial/FeH_logM.tex
Some things were updated, I guess

\documentclass[12pt, a4paper]{article}

\usepackage[english]{babel}
% Here the text area, the left margin and the top margin are set. 
% The right and bottom margins are automatically computed to fit the page. 
\usepackage{geometry}
\geometry{
    a4paper,  % 210 x 297 mm
    total={170mm,257mm},
    left=20mm,  % 210 - 170 = 40, so right margin will also be 20mm ?
    top=15mm,   % 297 - 257 = 40, so bottom margin will be 25mm ?
}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{float}
\usepackage{hyperref}
\hypersetup{pdfborder=0 0 0, colorlinks=true, linkcolor=blue, urlcolor=blue, citecolor=blue}
\usepackage{todonotes}
\usepackage{enumerate}
\usepackage{afterpage}    % Hack to remove page number from one specific page
\usepackage{transparent}  % Use transparent image to move float to right of two-col figure

\usepackage{natbib}
\usepackage{aas_macros}
\citestyle{aa}
\bibpunct{(}{)}{;}{a}{,}{,}

\graphicspath{{../img/}}   % Set image path

\renewcommand{\thefootnote}{\fnsymbol{footnote}}
\renewcommand*{\thesection}{4.\arabic{section}}
\setcounter{section}{-1}

\makeatletter
\renewcommand{\fnum@figure}{Fig. \thefigure}
\makeatother

\begin{document}
\part*{Sec. 4 \quad Results}
\textbf{Question.} Is the star formation model implemented for the Auriga simulations 
capable of producing a population of star particles that is consistent with the
observed properties of the Milky Way (MW) globular cluster system (GCS)? 
\begin{itemize}
    \item What data do we have available for MW and M31? \\ 
        \null\qquad $\rightarrow$ Sec.~\ref{sec:data} on p.~\pageref{sec:data}
    \item Does Auriga's star formation model produce sufficient star particles 
        with the right age and metallicity ([Fe/H]) to be consistent with the
        MW GCS? \\ 
        \null\qquad $\rightarrow$ Sec.~\ref{sec:FeH} on p.~\pageref{sec:FeH}
    \item Does Auriga's star formation model produce sufficient star particles
        with the right age and radial distribution to be consistent with the
        MW GCS? \\ 
        \null\qquad $\rightarrow$ Sec.~\ref{sec:Rgc} on p.~\pageref{sec:Rgc}
    \item Does Auriga's star formation model produce sufficient star particles
        with the right age, metallicity ([Fe/H]) and radial distribution to be
        consistent with the MW GCS? \\ 
        \null\qquad $\rightarrow$ Sec.~\ref{sec:FeH-Rgc} on p.~\pageref{sec:FeH-Rgc}
\end{itemize}
\textbf{Question.} What does the picture look like when broaden the scope to 
consider spirals in the Local Group (i.e. including Andromeda/M31)? \\

\noindent \textbf{Assumption I.} All star particles in the Auriga simulations with age
$> 10$~Gyr are globular cluster candidates. \\

\textbf{Justification} 
\begin{enumerate}[i)]
    {\setlength\itemindent{25pt} \item 
        \citet{2013ApJ...775..134V} measured [Fe/H] of 55 globular clusters in
        the MW and obtained age-estimates. The mean age of the MW GCS is 11.9~Gyr 
        with a dispersion of 0.9~Gyr. Furthermore, only one of the 55 GC 
        age-estimates is below 10~Gyr.
    }
    {\setlength\itemindent{25pt} \item 
        \citet{2017MNRAS.465.3622R} performed one simulation of a MW-like galaxy 
        (down to z=0.5, T$_{\rm lookback}\approx$~5~Gyr) and performs the entire
        analysis using a subset of star particles with ages $>10$~Gyr that is 
        referred to as `globular cluster candidates'.
    }
\end{enumerate}

\textbf{Weaknesses}
\begin{enumerate}[i)]
    {\setlength\itemindent{25pt} \item 
        \citet{2018MNRAS.475.4309P} show that star clusters do not simply follow 
        the same distribution as the field stars. In our approach we don't just 
        oversample the `real' population of globular clusters (by ignoring/excluding
        dynamical evolution), but the retrieved distributions of old star particles in 
        the simulations may not even faithfully represent the `true' distribution of 
        globulars.
    }
    {\setlength\itemindent{25pt} \item 
        \citet{2011AJ....141...61C} measured [Fe/H] of 87 globular clusters
        in M31 and obtained age-estimates. The mean age is 10.8~Gyr with a 
        dispersion of 2.3~Gyr. Furthermore, 27 GCs have age-estimates below 10~Gyr,
        with a minimum age-estimate of 4.8~Gyr. Perhaps an age cut of 6~Gyr 
        would would be more appropriate for M31, see Fig.~\ref{fig:MW-M31-age}.
    }
\end{enumerate}

\afterpage{%
\section{Available data for the Milky Way and Andromeda}
\label{sec:data}
\thispagestyle{empty}
\begin{figure*}[h!]
    \includegraphics[width=0.5\textwidth]{{MW_M31_Age_HistogramMassWeighted}.png}
    \includegraphics[width=0.49\textwidth]{{MW_M31_FeH_HistogramMassWeighted}.png}
    \caption{
        \emph{Left:} Mass-weighted age distribution of 55 GCs in the MW 
        \citep[data from][]{2013ApJ...775..134V} and 87 GCs 
        in M31 \citep[data from][]{2011AJ....141...61C}.
        \emph{Right:} Mass-weighted [Fe/H] distribution of 151 GCs in the MW 
        \citep[data from][2010 ed.]{1996AJ....112.1487H} and 314 GCs in M31.
        \label{fig:MW-M31-ageFeH}
    }

    \includegraphics[width=0.49\textwidth]{{MW_M31_Rgc_HistogramMassWeighted}.png}
    \includegraphics[width=0.53\textwidth]
        {{MW_M31_RgcFeH_HistogramMassWeighted_Harris1996ed2010data}.png}
    \caption{
        \emph{Left:} Mass-weighted R$_{\rm GC}$ distribution of 156 GCs in the MW.
        \emph{Right:} Mass-weighted [Fe/H]-R$_{\rm GC}$ distribution of
        151 GCs in the MW \citep[data from][2010 ed.]{1996AJ....112.1487H}, which
        is 98.19 \% of the total MW GCS mass.
        \label{fig:MW-M31-Rgc-FeHRgc}
    }

    \transparent{0.0}\includegraphics[width=0.49\textwidth]
        {{MW_M31_Rgc_HistogramMassWeighted}.png}
    \transparent{1.0}\includegraphics[width=0.53\textwidth]
        {{MW_M31_RgcFeH_HistogramMassWeighted_VandenBerg2013data}.png}
    \caption{
        Mass-weighted [Fe/H]-R$_{\rm GC}$ distribution of
        55 GCs in the Milky Way \citep[data from][]{2013ApJ...775..134V}, which
        is 43.27\% of the total MW GCS mass.
        \label{fig:MW-M31-FeHRgc2}
    }
\end{figure*}
\clearpage
}
\newpage

\afterpage{%
\newgeometry{
    a4paper,  % 210 x 297 mm
    total={170mm,257mm},
    left=20mm,  % 210 - 170 = 40, so right margin will also be 20mm ?
    top=5mm,   % 297 - 257 = 40, so bottom margin will be 25mm ?
}
\begin{table}[ht!]
    \centering
    \caption{
        The [Fe/H]-R$_{\rm GC}$ plot is generated using 
        \textsc{scipy.stats.binned\_statistic\_2d}. This table shows a manual 
        calculation of Msum and Ngc in each bin for debug purposes, just to check
        that \textsc{scipy}'s built-in method is used correctly. 
        Data from \citet[][2010 ed.]{1996AJ....112.1487H}.
        \label{tab:MW-FeH-Rgc}
    }
    \begin{tabular}{rrrrrrc}
        \hline
        xmin   & xmax   & ymin   & ymax   & Ngc    & Msum   & log$_{10}$(Msun) \\
        \hline
        1      & 3      & -2.5   & -1.5   & 7      & 1.7e+06 & 6.24       \\
        1      & 3      & -1.5   & -1.0   & 12     & 1.9e+06 & 6.27       \\
        1      & 3      & -1.0   & -0.5   & 9      & 7.7e+05 & 5.89       \\
        1      & 3      & -0.5   & -0.3   & 4      & 1.0e+06 & 6.02       \\
        1      & 3      & -0.3   & 0      & 2      & 3.2e+05 & 5.51       \\
        \hline
        3      & 8      & -2.5   & -1.5   & 16     & 4.6e+06 & 6.66       \\
        3      & 8      & -1.5   & -1.0   & 12     & 2.4e+06 & 6.37       \\
        3      & 8      & -1.0   & -0.5   & 10     & 2.5e+06 & 6.39       \\
        3      & 8      & -0.5   & -0.3   & 6      & 1.9e+06 & 6.29       \\
        3      & 8      & -0.3   & 0      & 1      & 3.0e+04 & 4.48       \\
        \hline
        8      & 15     & -2.5   & -1.5   & 11     & 3.0e+06 & 6.47       \\
        8      & 15     & -1.5   & -1.0   & 6      & 1.9e+06 & 6.27       \\
        8      & 15     & -1.0   & -0.5   & 1      & 6.5e+03 & 3.81       \\
        8      & 15     & -0.5   & -0.3   & 1      & 8.5e+04 & 4.93       \\
        8      & 15     & -0.3   & 0      & 0      & 0.0e+00 & -inf       \\
        \hline
        15     & 30     & -2.5   & -1.5   & 16     & 2.1e+06 & 6.31       \\
        15     & 30     & -1.5   & -1.0   & 6      & 2.2e+06 & 6.34       \\
        15     & 30     & -1.0   & -0.5   & 2      & 1.0e+04 & 4.02       \\
        15     & 30     & -0.5   & -0.3   & 1      & 1.5e+04 & 4.17       \\
        15     & 30     & -0.3   & 0      & 0      & 0.0e+00 & -inf       \\
        \hline
        30     & 125    & -2.5   & -1.5   & 6      & 1.1e+06 & 6.04       \\
        30     & 125    & -1.5   & -1.0   & 4      & 3.1e+05 & 5.49       \\
        30     & 125    & -1.0   & -0.5   & 1      & 1.4e+03 & 3.15       \\
        30     & 125    & -0.5   & -0.3   & 0      & 0.0e+00 & -inf       \\
        30     & 125    & -0.3   & 0      & 0      & 0.0e+00 & -inf       \\
        \hline
        \multicolumn{7}{c}{Data from \citet[][55 GCs]{2013ApJ...775..134V}} \\
        \hline
        1      & 3      & -2.5   & -1.5   & 3      & 5.5e+05 & 5.74       \\
        1      & 3      & -1.5   & -1.0   & 2      & 2.2e+05 & 5.35       \\
        1      & 3      & -1.0   & -0.5   & 2      & 2.3e+05 & 5.37       \\
        1      & 3      & -0.5   & -0.3   & 2      & 2.6e+05 & 5.42       \\
        1      & 3      & -0.3   & 0      & 0      & 0.0e+00 & -inf       \\
        \hline
        3      & 8      & -2.5   & -1.5   & 9      & 1.7e+06 & 6.22       \\
        3      & 8      & -1.5   & -1.0   & 5      & 9.1e+05 & 5.96       \\
        3      & 8      & -1.0   & -0.5   & 4      & 9.6e+05 & 5.98       \\
        3      & 8      & -0.5   & -0.3   & 1      & 1.1e+05 & 5.04       \\
        3      & 8      & -0.3   & 0      & 1      & 1.9e+05 & 5.29       \\
        \hline
        8      & 15     & -2.5   & -1.5   & 10     & 3.0e+06 & 6.48       \\
        8      & 15     & -1.5   & -1.0   & 4      & 1.3e+06 & 6.13       \\
        8      & 15     & -1.0   & -0.5   & 0      & 0.0e+00 & -inf       \\
        8      & 15     & -0.5   & -0.3   & 0      & 0.0e+00 & -inf       \\
        8      & 15     & -0.3   & 0      & 0      & 0.0e+00 & -inf       \\
        \hline
        15     & 30     & -2.5   & -1.5   & 6      & 6.8e+05 & 5.83       \\
        15     & 30     & -1.5   & -1.0   & 3      & 1.9e+06 & 6.29       \\
        15     & 30     & -1.0   & -0.5   & 1      & 8.9e+03 & 3.95       \\
        15     & 30     & -0.5   & -0.3   & 0      & 0.0e+00 & -inf       \\
        15     & 30     & -0.3   & 0      & 0      & 0.0e+00 & -inf       \\
        \hline
        30     & 125    & -2.5   & -1.5   & 0      & 0.0e+00 & -inf       \\
        30     & 125    & -1.5   & -1.0   & 0      & 0.0e+00 & -inf       \\
        30     & 125    & -1.0   & -0.5   & 0      & 0.0e+00 & -inf       \\
        30     & 125    & -0.5   & -0.3   & 0      & 0.0e+00 & -inf       \\
        30     & 125    & -0.3   & 0      & 0      & 0.0e+00 & -inf       \\
        \hline
    \end{tabular}
\end{table}
\clearpage
\restoregeometry
}
\clearpage

\section{Distribution of metallicity [Fe/H]}
\label{sec:FeH}

\section{Distribution of Galactocentric radii R$_{\rm GC}$}
\label{sec:Rgc}

\section{Distribution of [Fe/H]-R$_{\rm GC}$}
\label{sec:FeH-Rgc}

\newpage
\bibliographystyle{../mnras}
\bibliography{../AurigaGCS}

\end{document}